Methods Power Study

### 2022-08-12 simulation_study/simulate_cutoff.Rmd
updating with simulations outside the manuscript

---
title: "Simulate Cut Off"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

@@

```{r}
set.seed(895893)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(papaja)
```

# Introduction

This document outlines using the suggested procedure to simulate the number of participants necessary to accurately measure items. There are two key issues these ideas should address that we know about power:

1) we should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size)
2) we should see projected sample sizes that "level off" when pilot data increases. As with regular power estimates, studies can be "overpowered" to detect an effect, and this same idea should be present. For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.

# Explain the Procedure

1) Take your original pilot data.
2) Find a critical score for an accurately measured item using SE.
3) Sample from your pilot data using sample sizes starting at 20 participants and increase in units of 5 up to a value you feel is the maximum sample size.
4) Calculate the number of items that meet your critical score at each sample size.
5) Find the smallest sample size for 80%, 85%, 90%, and 95% power. 
6) Designate your minimum sample size from these values, the stopping rule used from step 2, and the maximum sample size from these values.

@@

*Population*. The data was simulated using the `rnorm` function assuming a normal distribution for 30 scale type items. Each population was simulated with 1000 data points. No items were rounded for this simulation.

First, the scale of the data was manipulated by creating three sets of scales. The first scale was mimicked after small rating scales (i.e., 1-7 type style) using a $\mu$ = 4 with a $\sigma$ = .25 around the mean to create item mean variability. The second scale included a larger potential distribution of scores with a $\mu$ = 50 ($\sigma$ = 10) imitating a 0-100 scale. Last, the final scale included a $\mu$ = 1000 ($\sigma$ = 150) simulating a study that may include response latency data in the milliseconds. While there are many potential scales, these three represent a large number of potential variables in the social sciences. As we are suggesting variances as a key factor for estimating sample sizes, the scale of the data is influential on the amount of *potential* variance. Smaller ranges of data (1-7) cannot necessarily have the same variance as larger ranges (0-100).

Next, item variance heterogeneity was included by manipulating the potential $\sigma$ for each individual item. For small scales, the $\sigma$ = 2 points with a variability of .2, .4, and .8 for low, medium, and high heterogeneity in the variances between items. For the medium scale of data, $\sigma$ = 25 with a variance of 4, 8, and 16. Last, for the large scale of data, $\sigma$ = 400 with a variance of 50, 100, and 200 for heterogeneity.

```{r}
# small potential variability overall, sort of 1-7ish scale
mu1 <- rnorm(30, 4, .25)
sigma1.1 <- rnorm(30, 2, .2)
sigma2.1 <- rnorm(30, 2, .4)
sigma3.1 <- rnorm(30, 2, .8)

@@

#evidence that they are simulated correctly
tapply(population1$score, list(population1$item, 
                               population1$scale), mean)
tapply(population1$score, list(population1$item,
                               population1$scale), sd)

tapply(population2$score, list(population2$item, 
                               population2$scale), mean)
tapply(population2$score, list(population2$item,
                               population2$scale), sd)

tapply(population3$score, list(population3$item, 
                               population3$scale), mean)
tapply(population3$score, list(population3$item,
                               population3$scale), sd) 
```

*Samples*. Each population was then sampled as if a researcher was conducting a pilot study. The sample sizes started at 20 participants per item increasing in units of 5 up to 100 participants.

```{r}
# create pilot samples from 20 to 100
samples1 <- samples2 <- samples3 <- list() 
sizes <- c(20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90, 95, 100)
for (i in 1:length(sizes)){
  samples1[[i]] <- population1 %>% 
    group_by(item, scale) %>% 
    slice_sample(n = sizes[i])

@@

  samples3[[i]] <- population3 %>% 
    group_by(item, scale) %>% 
    slice_sample(n = sizes[i])
}
```

*AIPE Criterions*. The standard errors of each item were calculated to mimic the AIPE procedure of finding an appropriately small confidence interval, as standard error functions as the main component in the formula for normal distribution confidence intervals. Standard errors were calculated at each decile of the items up to 90% (0% smallest SE, 10% ..., 90% largest SE). The lower deciles would represent a strict criterion for accurate measurement, as many items would need smaller SEs to meet AIPE goals, while the higher deciles would represent less strict criterions for AIPE goals.

```{r}
# calculate the SEs and the cutoff scores 
SES1 <- SES2 <- SES3 <- list()
cutoffs1 <- cutoffs2 <- cutoffs3 <- list()
sd_items1 <- sd_items2 <- sd_items3 <- list()
for (i in 1:length(samples1)){
  sd_items1[[i]] <- samples1[[i]] %>% group_by(item, scale) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
  sd_items2[[i]] <- samples2[[i]] %>% group_by(item, scale) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
  sd_items3[[i]] <- samples3[[i]] %>% group_by(item, scale) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))

@@

  cutoffs1[[i]] <- apply(as.data.frame(SES1[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1))
  cutoffs2[[i]] <- apply(as.data.frame(SES2[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1))
  cutoffs3[[i]] <- apply(as.data.frame(SES3[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1))
}
```

## AIPE Simulation

In this section, we simulate what a researcher might do if they follow our suggested application of AIPE to sample size planning based on well measured items. Assuming each pilot sample represents a dataset a researcher has collected, we will simulate samples of 20 to 500 to determine what the new sample size suggestion would be. We assume that samples over 500 may be considered too large for many researchers who do not work in teams or have participant funds. The standard error of each item was calculated for each suggested sample size by pilot sample size by population type.

```{r}
# sequence of sample sizes to try
samplesize_values <- seq(20, 500, 5)

@@

} # end all sample loop 
```

Next, the percent of items that fall below the cutoff scores, and thus, would be considered "well-measured" were calculated for each decile by sample.

```{r}
summary_list1 <- summary_list2 <- summary_list3 <- list()
for (i in 1:length(sampled_values1)){

@@

summary_DF <- bind_rows(summary_list1, 
                        summary_list2, 
                        summary_list3)
```

From this data, we pinpoint the smallest suggested sample size at which 80%, 85%, 90%, and 95% of the items fall below the cutoff criterion. These values were chosen as popular measures of "power" in which one could determine the minimum suggested sample size (potentially 80% of items) and the maximum suggested sample size (potentially 90%).

```{r}
summary_long_80 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
  filter(value >= .80) %>% 
  arrange(sample_size, original_n, source, scale, name) %>% 
  group_by(original_n, name, source, scale) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 80)

summary_long_85 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
  filter(value >= .85) %>% 
  arrange(sample_size, original_n, source, scale, name) %>% 
  group_by(original_n, name, source, scale) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 85)

summary_long_90 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
  filter(value >= .90) %>% 
  arrange(sample_size, original_n, source, scale, name) %>% 
  group_by(original_n, name, source, scale) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 90)

summary_long_95 <- summary_DF %>% 
  pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
  filter(value >= .95) %>% 
  arrange(sample_size, original_n, source, scale, name) %>% 
  group_by(original_n, name, source, scale) %>% 
  slice_head(n = 1) %>% 
  mutate(power = 95)

summary_long <- rbind(summary_long_80, 
                      summary_long_85,
                      summary_long_90,
                      summary_long_95)

summary_long$source <- factor(summary_long$source, 
                              levels = c("low", "med", "high"),
                              labels = c("Low Variance", "Medium Variance", "High Variance"))

@@

# Results

## Differences in Item Variance

We examined if this procedure is sensitive to differences in item heterogeneity, as we should expect to collect larger samples if we wish to have a large number of items reach a threshold of acceptable variance; potentially, assuring we *could* average them if a researcher did not wish to use a more complex analysis such as multilevel modeling.

The figure below illustrates the potential minimum sample size for 80% of items to achieve a desired cutoff score. The black dots denote the original sample size against the suggested sample size. By comparing the facets, we can determine that our suggested procedure does capture the differences in heterogeneity. As heterogeneity increases in item variances, the proposed sample size also increases, especially at stricter cutoffs. Missing cutoff points where sample sizes proposed would be higher than 500.

```{r}
summary_long$source <- factor(summary_long$source, 
                              levels = c("Low Variance", "Medium Variance", "High Variance"))

ggplot(summary_long %>% filter(power == 80), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ source*scale2) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

## Sample Size Sensitivity to Pilot Data Size

In our second question, we examined if the suggested procedure was sensitive to the amount of information present in the pilot data. Larger pilot data is more informative, and therefore, we should expect a lower projected sample size. As shown in the figure below for only the low variability and small scale data, we do not find this effect. These simplistic simulations from the pilot data would nearly always suggest a larger sample size - mostly in a linear trend increasing with sample sizes. This result comes from the nature of the procedure - if we base our estimates on some SE cutoff, we will almost always need a bit more people for items to meet those goals. This result does not achieve our second goal.

```{r}
ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

Therefore, we suggest using a correction factor on the simulation procedure to account for the known asymptotic nature of power (i.e., at larger sample sizes power increases level off). For this function, we combined a correction factor for upward biasing of effect sizes (Hedges' correction) with the formula for exponential decay calculations. The decay factor is calculated as follows:

$$ 1 - \sqrt{\frac{N_{pilot} - min(N_{sim})}{N_{pilot}}}^{log_2(N_{pilot})}$$
$N_{pilot}$ indicates the sample size of the pilot data minus the minimum sample size for simulation to ensure that the smallest sample sizes do not decay (e.g., the formula zeroes out). This value is raised to the power of $log_2$ of the sample size of the pilot data, which decreases the impact of the decay to smaller increments for increasing sample sizes. This value is then multiplied by the proposed sample size. As show in the figure below, this correction factor produces the desired quality of maintaining that small pilot studies should *increase* sample size, and that sample size suggestions level off as pilot study data sample size increases.

```{r}
decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)

summary_long$new_sample <- summary_long$sample_size*decay

ggplot(summary_long %>% filter(source == "Low Variance") %>% filter(scale == 1), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

## Corrections for Individual Researchers

We have portrayed that this procedure, with a correction factor, can perform as desired. However, within real scenarios, researchers will only have one pilot sample, not the various simulated samples shown above. What should the researcher do to correct their sample size on their own pilot data?

To explore if we could recover the new suggested sample size from data a researcher would have, we used linear models to create a formula for calculation. First, the corrected sample size was predicted by the original suggested sample size. Next, the standard deviation of the item standard deviations was added to the equation to recreate heterogeneity estimates. Last, we included the pilot sample size. The scale of the data is embedded into the standard deviation of the items (*r* = `r round(cor(summary_long$scale, summary_long$sd_item), digits = 2)`), and therefore, this variable was not included separately. The standard deviation of the item's standard deviation embeds both heterogeneity and potential scale size.

```{r}
user_model <- lm(new_sample ~ sample_size, data = summary_long)
user_print <- apa_print(user_model)

user_model2 <- lm(new_sample ~ sample_size + sd_item, data = summary_long)
user_print2 <- apa_print(user_model2)

user_model3 <- lm(new_sample ~ sample_size + sd_item + original_n, data = summary_long)
user_print3 <- apa_print(user_model3)

change_table <- tidy(anova(user_model, user_model2, user_model3))
```

The first model using original sample size to predict new sample size was significant, `r user_print$full_result$modelfit$r2`, capturing nearly 90% of the variance, `r user_print$estimate$sample_size`. The second model with item standard deviation was better then the first model *F*(`r change_table$df[2]`, `r change_table$res.df[2]`) = `r round(change_table$statistic[2], digits = 2)`, *p* < .001, `r user_print2$estimate$modelfit$r2`. The item standard deviation predictor was significant, `r user_print2$full_result$sd_item`. The addition of the original pilot sample size was also significant, *F*(`r change_table$df[3]`, `r change_table$res.df[3]`) = `r round(change_table$statistic[3], digits = 2)`, *p* < .001, `r user_print3$estimate$modelfit$r2`.

As shown in the table below for our final model, the new suggested sample size is proportional to the original suggested sample size (i.e., *b* < 1), which reduces the sample size suggestion. As variability increases, the suggested sample size also increases to capture differences in heterogeneity shown above; however, this predictor is not significant in the final model, and only contributes a small portion of overall variance. Last, in order to correct for large pilot data, the original pilot sample size decreases the new suggested sample size. This formula approximation captures 96% of the variance in sample size scores and should allow a researcher to estimate based on their own data.

```{r results = 'asis'}
apa_table(tidy(user_model3))
```

## Choosing an Appropriate Cutoff

```{r}
by_cutoff <- list()
R2 <- list()

for (cutoff in unique(summary_long$name)){
  by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + sd_item + original_n, data = summary_long  %>% filter(name == cutoff))
  R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
}

R2
```

Last, we examine the question of an appropriate SE decile. All graphs for power, heterogeneity, scale, and correction are presented below. First, the 0%, 10%, and 20% deciles are likely too restrictive, providing very large estimates that do not always find a reasonable sample size in proportion to the pilot sample size, scale, and heterogeneity.If we examine the $R^2$ values for each decile of our regression equation separately, we find that the 50% (`r round(R2[["Decile 50"]], digits = 3)`) represents the best match to our corrected sample size suggestions. The 50% decile, in the corrected format, appears to meet all goals: 1) increases with heterogeneity and scale of data, and 2) higher suggested values for small original samples and a leveling effect at larger pilot data.

The formula for finding the corrected sample size using a 50% decile is: $Final Sample = 39.269 + 0.700 \times X_{Suggested Sample Size} + 0.003 \times X_{SD Items} + -0.695 \times X_{Pilot Sample Size}$. The suggested sample size will be estimated from the 80%, 85%, 90%, or 95% selection at the 50% decile as shown above. The item SD can be calculated directly from the data, and the pilot sample size is the sample size of the data from which a researcher is simulating their samples.

## Conclusions

1) Take your original pilot data
2) Find a critical score for an accurately measured item using SE _*at the 50% decile of the item SE values*_.
3) Sample from your pilot data using sample sizes starting at 20 participants and increase in units of 5 up to a value you feel is the maximum sample size. 
4) Calculate the number of items that meet your critical score at each sample size.
5) Find the smallest sample size for 80%, 85%, 90%, and 95% power. 
6) _*Apply the correction to the smallest sample sizes*_.
7) Designate your minimum sample size from these values, the stopping rule used from step 2, and the maximum sample size from these values.

### Low Variability

```{r}
ggplot(summary_long %>% filter(source == "Low Variance"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ scale2*power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

```{r}
ggplot(summary_long %>% filter(source == "Low Variance"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ scale2*power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

### Medium Varability

```{r}
ggplot(summary_long %>% filter(source == "Medium Variance"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ scale2*power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

```{r}
summary_long$new_sample <- summary_long$sample_size*decay

ggplot(summary_long %>% filter(source == "Medium Variance"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ scale2*power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

### High Variability

```{r}
ggplot(summary_long %>% filter(source == "High Variance"), aes(original_n, sample_size, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ scale2*power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

```{r}
ggplot(summary_long %>% filter(source == "High Variance"), aes(original_n, new_sample, color = name)) + 
  geom_point() +  
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  theme_classic() + 
  facet_wrap(~ scale2*power) + 
  xlab("Original Sample Size") + 
  ylab("Suggested Sample Size") + 
  scale_color_discrete(name = "Cutoff Score")
```

>>>>

---
title: "Methods Power Study"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

@@

```{r}
set.seed(895893)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(papaja)
library(rio)
```

This document includes the informal write up for the method for the manuscript. We exported the simulations to this document in order to run the simulations separately from the manuscript for speed reasons (i.e., it's really slow to run the simulations along with the manuscript markdown). At the end of this document, a `simulated_summary_data.csv` file is created that is imported into the manuscript for plotting and other analyses. Please see the manuscript for formal write up.

@@

*Population*.

- Simulate the data with `rnorm` assuming a normal distribution for 30 scale items
- Population is created with 1000 data points 
- No rounding 
- We will run this part 100 times, so create a function that does this for you.

```{r sim_pop, include = T, echo = T}
simulate_population <- function (x = 1){

  # small potential variability overall, sort of 1-7ish scale
  mu1 <- rnorm(30, 4, .25)
  sigma1.1 <- rnorm(30, 2, .2)
  sigma2.1 <- rnorm(30, 2, .4)
  sigma3.1 <- rnorm(30, 2, .8)

@@

  # return populations 
  return(list(population1 = population1, 
              population2 = population2, 
              population3 = population3))
}
```

- Within that function, create the scale size variable: 
  - Small scale data 1-7, mean 4, sd of mean .25
  - Medium scale data 0-100, mean 50, sd of mean 10
  - Large scale data no range, mean 1000, sd of mean 150
  - These choices are meant to mimic popular scale choices 
- Additionally, create the scale heterogeneity variable:
  - Small scale heterogeneity, sd 2, sd of sd .2 .4 .8
  - Medium scale heterogeneity, sd 25, sd of sd 4, 8, 16
  - Large scale heterogeneity, sd 400, sd of sd 50, 100, 200

*Samples*.

- Create a function that does the samples that a researcher might do
- Includes a start and stop of the number of participants you might consider 
- Includes an increasing value between samples
- We will also run this part 100 times

```{r sim_sample, include = T, echo = T}
simulate_samples <- function(start = 20, stop = 100, increase = 5,
                             population1, population2, population3){

  # save those samples for small medium large 
  samples1 <- samples2 <- samples3 <- list()

  # create the list of sizes 
  sizes <- seq(from = start, to = stop, by = increase)

  # loop over sizes and create those samples 
  for (i in 1:length(sizes)){
    samples1[[i]] <- population1 %>% 
      group_by(item, scale) %>% 
      slice_sample(n = sizes[i])

@@

    samples3[[i]] <- population3 %>%
      group_by(item, scale) %>%
      slice_sample(n = sizes[i])
  }

  # return your values
  return(list(samples1 = samples1, 
              samples2 = samples2, 
              samples3 = samples3,
              sizes = sizes))

}
```

*Cutoff Score Criterions*.

- Calculate the SEs for the items at each decile 
- Save as a function for later

```{r calc_SE, include = T, echo = T}
calculate_deciles <- function(samples1, samples2, samples3){
  # calculate the SEs and the cutoff scores 
  SES1 <- SES2 <- SES3 <- list()
  cutoffs1 <- cutoffs2 <- cutoffs3 <- list()
  sd_items1 <- sd_items2 <- sd_items3 <- list()

  # loop and calculate for each sample size 
  for (i in 1:length(samples1)){
  sd_items1[[i]] <- samples1[[i]] %>% group_by(item, scale) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
  sd_items2[[i]] <- samples2[[i]] %>% group_by(item, scale) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))
  sd_items3[[i]] <- samples3[[i]] %>% group_by(item, scale) %>% 
    summarize(sd = sd(score), .groups = "keep") %>% 
    ungroup() %>% group_by(scale) %>% summarize(sd_item = sd(sd))

@@

  cutoffs1[[i]] <- apply(as.data.frame(SES1[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1),
                         na.rm = T)
  cutoffs2[[i]] <- apply(as.data.frame(SES2[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1),
                         na.rm = T)
  cutoffs3[[i]] <- apply(as.data.frame(SES3[[i]]), 2, 
                         quantile, 
                         probs = seq(0, .9, by = .1),
                         na.rm = T)

  # return information
  return(list(SES1 = SES1, SES2 = SES2, SES3 = SES3, 
              cutoffs1 = cutoffs1, cutoffs2 = cutoffs2, cutoffs3 = cutoffs3, 
              sd_items1 = sd_items1, sd_items2 = sd_items2, sd_items3 = sd_items3))

}
```

## Researcher Sample Simulation

- Here we are simulating the researcher side of the equation assuming they have a sample pulled from population (which we simulated above)
- We will simulate samples from 20 to 2000 ... mostly only to see the leveling out effect of the simulation procedure, as we likely believe that over 2000 is not reasonable for most researchers. 
- Calculate the SEs here, to see if it matches our cutoff 
- Save this as a function to run within our pipeline (100 populations and 100 samples)

```{r sim_sim, include = T, echo = T}
simulate_researcher <- function(samples1, samples2, samples3, 
                                start = 20, stop = 2000, 
                                increase = 5){

  # sequence of sample sizes to try
  samplesize_values <- seq(from = start, to = stop, by = increase)

@@

  } # end all sample loop

  # return sampled values 
  return(list(sampled_values1 = sampled_values1, 
              sampled_values2 = sampled_values2, 
              sampled_values3 = sampled_values3,
              samplesize_values = samplesize_values))

}
```

- Next, calculate the percent of items falling below the decile scores 
- Pick up 80, 85, 90, and 95% of items to mimic power 
- Save a function for our pipeline

```{r calc_percent, include = T, echo = T}
calculate_percent <- function(sampled_values1, sampled_values2, sampled_values3,
                              cutoffs1, cutoffs2, cutoffs3, 
                              samplesize_values, sizes){

  # create temporary storage
  summary_list1 <- summary_list2 <- summary_list3 <- list()

  # loop and calculate
  for (i in 1:length(sampled_values1)){

@@

    } # end loop and calculate

  # create end summary 
  summary_DF <- bind_rows(summary_list1, 
                        summary_list2, 
                        summary_list3)

  # return values 
  return(summary_DF)

}
```

- Last combine everything together grabbing the proposed sample at each decile given the 80, 85, 90, and 95 scores 
- Create a function for our pipeline

```{r calc_percent_combine, include = T, echo = T}
calculate_proposed <- function(summary_DF){

  # 80% summary 
  summary_long_80 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
    filter(value >= .80) %>% 
    arrange(sample_size, original_n, source, scale, name) %>% 
    group_by(original_n, name, source, scale) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 80)

  # 85% summary
  summary_long_85 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
    filter(value >= .85) %>% 
    arrange(sample_size, original_n, source, scale, name) %>% 
    group_by(original_n, name, source, scale) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 85)

  # 90% summary  
  summary_long_90 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
    filter(value >= .90) %>% 
    arrange(sample_size, original_n, source, scale, name) %>% 
    group_by(original_n, name, source, scale) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 90)

  # 95% summary
  summary_long_95 <- summary_DF %>% 
    pivot_longer(cols = -c(sample_size, original_n, source, scale)) %>% 
    filter(value >= .95) %>% 
    arrange(sample_size, original_n, source, scale, name) %>% 
    group_by(original_n, name, source, scale) %>% 
    slice_head(n = 1) %>% 
    mutate(power = 95)

  # combine the summary together
  summary_long <- rbind(summary_long_80, 
                        summary_long_85,
                        summary_long_90,
                        summary_long_95)

  # return values
  return(summary_long)

}
```

## Simulation Pipeline

- Put together all the simulation pieces 
  - Simulate the population 
  - Simulate samples within that population 
  - Calculate the cutoffs and deciles 
  - Simulate like the researcher would 
  - Calculate the percent under the decile
  - Create final summary

- We will simulate multiple times at each simulation point:
  - Populations X 100
  - Samples X 100
  - Simulate Researcher X 100

```{r simulation_pipeline, include = T, echo = T}
# a list to save this in 
full_simulation <- list()

for (i in 1:10){

  # simulate population 
  populations <- simulate_population(x)

  for (p in 1:10){

    # simulate samples from population
    samples <- simulate_samples(start = 20, stop = 100, increase = 5, 
                            population1 = populations$population1,
                            population2 = populations$population2, 
                            population3 = populations$population3)

    # calculate deciles 
    deciles <- calculate_deciles(samples1 = samples$samples1, 
                             samples2 = samples$samples2,
                             samples3 = samples$samples3)

    # reset this with each new sample 
    researcher_summary <- list()

    for (q in 1:10){

      # simulate researcher
      researcher_data <- simulate_researcher(samples1 = samples$samples1, 
                                       samples2 = samples$samples2,
                                       samples3 = samples$samples3, 
                                       start = 20, stop = 30, 
                                       increase = 5)

      # calculate percent under each decile
      percent_items <- calculate_percent(
        sampled_values1 = researcher_data$sampled_values1, 
        sampled_values2 = researcher_data$sampled_values2, 
        sampled_values3 = researcher_data$sampled_values3, 
        cutoffs1 = deciles$cutoffs1, 
        cutoffs2 = deciles$cutoffs2, 
        cutoffs3 = deciles$cutoffs3, 
        samplesize_values = researcher_data$samplesize_values,
        sizes = samples$sizes)

      # calculate final summary data 
      researcher_summary[[q]] <- calculate_proposed(summary_DF = percent_items)

    } # tell the researcher to do it 1000 times for normal simulation numbers

    # after the researcher summary, summarize and stick in dataframe 
    # so we have a dataframe of each population by sample 
    full_simulation[[paste(i, "_", q, sep = "")]] <- 
      bind_rows(researcher_summary) %>% 
      group_by(scale, original_n, source, name, power) %>% 
      summarize(sample_size = mean(sample_size),
                value = mean(value), .groups = "keep")

  } # simulate the sample 100 times for each population

} # simulate the population 100 times

```

```{r}
# clean up the stuff from the summary 
  summary_long$source <- factor(summary_long$source, 
                                levels = c("low", "med", "high"),
                                labels = c("Low Variance", "Medium Variance", "High Variance"))

@@
